Monitoring and Projecting Global Hunger: Are We On Track?

### 2021-06-11 docs/response_to_reviewers.md

---
created: 2021-06-10T10:04:08-0400
modified: 2021-06-11T12:16:25-0400
title: Response to Reviewers
header-includes:
 - \hypersetup{colorlinks=true}
 - \hypersetup{urlcolor=blue}
---

This paper presents a global assessment of food security outcomes and prediction for the years to come. It leverages a useful household-level food security dataset and gap fills with a random forest-model to predict food security outcomes in missing locations. While it has potential to add important findings to the literature, there are several methodological gaps that need to be addressed before it can be considered for publication. The General Comments below reflect broader specific locations in the text where I take issue. That said, I have not fully addressed line-by-line critiques yet because the broader concerns are too numerousR I highly encourage the authors to not be discouraged, as I see this work as immensely valuable, but I want them to produce the most robust science possible.

First, there is a growing body of literature that shows wide variation estimates of population across gridded gridded population products (e.g., Chen et al. 2020; Bustos et al. 2020; Tuholske et al. 2019) and recently published fitness-for-use guidelines (see Leyk et al. 2019). The authors do not clearly explain why they chose WorldPop (as opposed to Landscan, ESRI's World Population Estimates, the Global Human Settlement Layer, High Resolution - see www.popgrid.org for more). Furthermore, it is not clear how the Worldpop data is even used. It is mentioned in the Main Text, but I do not see it in your Supplemental Materials. Population estimates should be re-run using other gridded population datasets as I suspect that they will yield markedly different results.

**We regret that we did not make it more clear how the population data was used - we have now made that more clear in the text, and that may have also made it more apparent why we chose the WorldPop dataset.  We use gridded population data to estimate demographic characteristics for administrative areas, and chose to use WorldPop because it has data on age and sex.  Examining the other datasets you mention and searching through that very useful review by Leyk et al., it seems that GPWv4 also includes such demographic data, but more coarsely aggregated (age categories are across wide ranges, such as 15-49, whereas WorldPop bins ages every five years).  Moreover, WorldPop is an annual product that we can match to the specific year of our survey data, whereas GPWv4 is only available for 2015.  We have now made it more apparent in the _Disaggregation_ subsection of the _Methods_ section why we chose WorldPop, and have added some discussion and relevant citations from your review.**

**Additionally, we also used population data at the administrative area level to estimate the food insecure population totals after modeling rates of food insecurity.  This data came from the [Subnational Human Development Database (SHDI)](https://doi.org/10.1038/sdata.2019.38), which in turn is derived from national censuses.  Because this analysis is conducted at the admin-area level, we believe it is not necessary to use high-resolution gridded data to estimate population totals and the SHDI data is sufficient.**

Second, like the population data employed, I have strong concerns with the temperature and precipitation data and methodology employed here. Why use TerraClimate over the widely used and higher resolution CHIRPS (Funk et al. 2015) or other climate reanalysis (ERA5, MERRA-2, etc.) and the new CHIRTS-Max (Funk et al 2019) or CHIRTS-daily (Verdin et al. 2020) high-resolution temperature products to establish your baseline? What temporal resolution is 'mean temperature'? Surely it is widely established the annualized climate data over large distance data does not reflect the true nature of coupling between household food security outcomes and climate (see Shukla et al. 2020 for one recent example). I point to the example of Sternberg 2012 that clearly shows how climate shocks/variability across the planet can telescope to food security outcomes at great distances. How climate (across spatial and temporal scales) affect local price and labor is key here, yet you do not even
discuss price at all, much less employ it, in your modeling paradigm. Yet we know that specifically for urban areas, where half the planet lives, food security outcomes are highly dependent on labor and price (see Blekking et al. 2020 for an example). Also, what about seasonality (again see Shukla et al. 2020 for one example)? None of these crucial aspects of household-level food security outcomes are clearly addressed in the current version of the manuscript. I suspect that the reason your 'climate' variables do not predict your outcome is because you spatially and temporally smooth your data to the point where it will no matter what not capture the true nature of how climate shocks impact food security (e.g. huge spatiotemporal coupling between where food is grown, when it is grown, how much the price is dictated by local, national, and global market forces, etc.). See Niles et al. 2021 for a much more robust approach.

**We thank you for this detailed feedback, and we agree that the approach to modeling the impacts of climate on food security is weak.  However, we are in many ways hobbled by our survey data, which is only available at the annual temporal scale and only reported at the national spatial scale, which we use to infer subnational scales.  This makes it much harder to trace the impacts of climate on food security and necessitates the smoothing of our climate data.  The papers you cite by Niles, Blekking, and Shukla all seem to have indicators of food security available at fine spatial scales (Niles uses DHS clusters/community GPS points, Blekking uses household GPS points) and fine temporal scales (Niles uses monthly data, Shukla uses Dekads).**

**We also note that it is important to consider the scope of our analysis.  Weather shocks as well as labor and food prices certainly explain a lot of the variation in food security between households and across seasons.  However, the scope of this analysis is to explore drivers of food insecurity across the globe and over decadal timeframes.  We are especially interested in modeling future trends in food security.  We therefore use GCM projections, which are useful insofar as they are indicative of shifting baselines, but are unable to anticipate actual droughts or floods that would impact food insecurity in 2030.**

**Given these considerations of how the scale of our data (annual, admin-area scale) and scope of our analysis (modeling food insecurity globally and out to 2030) makes it infeasible to explore the effect of climate on food insecurity beyond shifting baselines, we do not think that using other datasets like ERA-5 or MERRA-2 would add much value. (Also, the CHIRPS and CHIRTS datasets, while excellent, are not truly global and unavailable for certainly latitudes, so couldnt be applied to this analysis).  However, you raise very important considerations in your feedback and we therefore make two improvements to the paper: (1) in addition to the survey year's rainfall totals and precipitation averages, we add a feature for the previous year's total rainfall and mean temperature, and additionally (2) we add a paragraph to the discussion highlighting many of the points you raise and further outlining the considerations that led to the analytical approach we took.**

Third, Random Forrest models suffer from training data bias. If the underlying training data if flawed, the Random Forest Model has no way of identifying this. What quality control is used to determine if your survey data is correct? One way to address this is to compare the Gallup data with other sources like DHS or LSMS surveys from the same year. Do they both measure similar national-level food security outcomes? I am also curious to know how your models preform with simple linear or logistic regression? Do you get the same results? I recognize Random Forest are nonparametric, but I would be keen to see how linear models preform with controls. Also, what happens when you trade out wasting/stunting as your predicted outcome versus the FIES-based metric?

Fourth, the spatial nature of your modeling is confusing and not clear. Are metrics averaged at the sub-national scale (if so what level of analysis and for which country)? How is degree of urbanization quantified (e.g. are Worldpop pixels masked at some threshold?)? Is the Gallup representative at the national level? District-level? What controls are used to ensure that the Gallup data is a representative sample (Not even explained in the Supplemental Materials)? Your methodology (including how you address issues of spatial scale) needs to be more clearly described in the Main Text and the Supplemental Material before this paper can be considered for publication.

https://userforum.dhsprogram.com/index.php?t=msg&th=7832&start=0&

Fifth, I take issue with this being the first global food security paper at a sub-national level (Pg 12, ln 11). Niles et al. is another similar paper, that is near global, with a much more robust modeling framework. Additionally, FEWS-net and WFP provide these assessments regularly or real time. Specifically, you do not present any sub-national findings. What are they? Why is sub-national important and how do your results advance our understanding of food security over space and time? Surely there may be important nuances in the sub-national data (e.g. Does Nairobi look like Dhaka? Is your assumption that two households with similar characteristics, irrespective of geography, correct? Do your results show that this assumption is correct?).

Finally, and importantly, how are you defining 'urban'? (See Catteneo et al. 2021 for further details). This is not clear in your Supplementary Material or main text. You do not address the growing body of literature that clearly demonstrates important differences between 'rural' and 'urban' food security measurement and outcomes (see Blekking et al. 2020, Tuholske et al. 2020, Haysom & Tawodzera 2018, Cockx et al. 2018) that are extremely important for stakeholders and policy development (Battersby 2017) and poverty alleviation (Christiaensen et al. 2017).

A few specifics:

Pg. 3, Ln 35 - 38: This assumption needs strong justification. Recent urban food security studies show wide heterogeneity in food security outcomes among 'low-income' households depending on the food security metric used (see Blekking et al. 2020, Tuholske et al. 2020, etc.). 'Urbanization level' at an administrative-unit level (if that is what you are using…it's not clear), thus, may not be good predictor of food security outcomes.

Pg. 4, Ln 20 - 23: Please expand your justification as to why you selected these climate datasets. Why is spatial averaging over large length scales appropriate to predict food security outcomes? Why is annual average precipitation and (I assume annual) average temperature appropriate?

Pg. 7, Ln 19 - 30: This results is in contrast to FEWS NET outlook that the those at risk to famine rose by 85 million people from 2015 - 2019 (I can't find the brief, but here is an article with the numbers: https://www.sciencemag.org/news/2020/04/how-team-scientists-studying-drought-helped-build-world-s-leading-famine-prediction). This was before cofid. Why do your numbers differ?

Pg. 9, 3.2: You have sub-national data, and you make assumptions that the characteristics of two similar households have the same food security outcomes, yet you present your findings at national level. An increasing body of research (e.g. ref) is moving away from national-level estimates as they mask the true nature of food security outcomes and are not relevant for policy maker or targeted interventions. Please justify these findings given the high degree of spatiotemporal heterogeneity in food security outcomes.

References

Battersby, J. (2017). MDGs to SDGs-new goals, same gaps: the continued absence of urban food security in the post-2015 global development agenda. African Geographical Review, 36(1), 115-129.

Blekking, J., Waldman, K., Tuholske, C., & Evans, T. (2020). Formal/informal employment and urban food security in Sub-Saharan Africa. Applied Geography, 114, 102131.

Bustos, M. F. A., Hall, O., Niedomysl, T., & Ernstson, U. (2020). A pixel level evaluation of five multitemporal global gridded population datasets: a case study in sweden, 1990-2015. Population and environment, 1-23.

Cattaneo, A., Nelson, A., & McMenomy, T. (2021). Global mapping of urban-rural catchment areas reveals unequal access to services. Proceedings of the National Academy of Sciences, 118(2).

Chen R, Yan H, Liu F, Du W, Yang Y. Multiple Global Population Datasets: Differences and Spatial Distribution Characteristics. ISPRS International Journal of Geo-Information. 2020;9: 637.

Christiaensen, L., & Kanbur, R. (2017). Secondary towns and poverty reduction: refocusing the urbanization agenda. Annual Review of Resource Economics, 9, 405-419.

Cockx, L., Colen, L., & De Weerdt, J. (2018). From corn to popcorn? Urbanization and dietary change: evidence from rural-urban migrants in Tanzania. World Development, 110, 140-159.

Funk, C., Peterson, P., Landsfeld, M., Pedreros, D., Verdin, J., Shukla, S., ... & Michaelsen, J. (2015). The climate hazards infrared precipitation with stations—a new environmental record for monitoring extremes. Scientific data, 2(1), 1-21.

Funk, C., Peterson, P., Peterson, S., Shukla, S., Davenport, F., Michaelsen, J., ... & Mata, N. (2019). A high-resolution 1983-2016 T max climate data record based on infrared temperatures and stations by the climate Hazard center. Journal of Climate, 32(17), 5639-5658.
Haysom, G., & Tawodzera, G. (2018). "Measurement drives diagnosis and response": Gaps in transferring food security assessment to the urban scale. Food Policy, 74, 117-125.

Leyk, S., Gaughan, A. E., Adamo, S. B., Sherbinin, A. D., Balk, D., Freire, S., ... & Pesaresi, M. (2019). The spatial allocation of population: a review of large-scale gridded population data products and their fitness for use. Earth System Science Data, 11(3), 1385-1409.

Niles, M. T., Emery, B. F., Wiltshire, S., Brown, M. E., Fisher, B., & Ricketts, T. H. (2021). Climate impacts associated with reduced diet diversity in children across nineteen countries. Environmental Research Letters, 16(1), 015010.

Shukla, S., Husak, G., Turner, W., Davenport, F., Funk, C., Harrison, L., & Krell, N. (2021). A slow rainy season onset is a reliable harbinger of drought in most food insecure regions in Sub-Saharan Africa. Plos one, 16(1), e0242883.

Sternberg, T. (2012). Chinese drought, bread and the Arab Spring. Applied Geography, 34, 519-524.

Tuholske, C., Caylor, K., Evans, T., & Avery, R. (2019). Variability in urban population distributions across Africa. Environmental Research Letters, 14(8), 085009.

Tuholske, C., Andam, K., Blekking, J., Evans, T., & Caylor, K. (2020). Comparing measures of urban food security in Accra, Ghana. Food Security, 1-15.

Verdin, A., Funk, C., Peterson, P., Landsfeld, M., Tuholske, C., & Grace, K. (2020). Development and validation of the CHIRTS-daily quasi-global high-resolution daily temperature data set. Scientific Data, 7(1), 1-14.

### 2021-06-11 docs/modeling_fies_update.tex
being making edits for reviewer

%\maketitle
\begin{abstract}

This paper presents the first global picture of food security at a subnational level based on the Food Insecurity Experience Scale, the indicator for the Sustainable Development Goal of ``Zero Hunger" that is most indicative of the individual's lived experience of food insecurity and hunger.  Using microdata from 75 countries and filling gaps using machine learning, we find significant heterogeneity in levels of food insecurity around the world.  Examining global temporal trends and accounting for the effects of the COVID-19 pandemic, we find that rates of severe food insecurity are declining, resulting in global decreases in the total number of severely food insecure people.  However, the total number of moderately food insecure people has been increasing and, after recovering from the shock of the COVID-19 pandemic, we predict it will continue to increase through the end of the 2020s.  Overall, we conclude that current trends in development and demographic change will still leave a large share of the world's population still experiencing hunger by 2030.  
\end{abstract}

@@

This paper contributes to shifting the research frontier by filling these two gaps in terms of spatial coverage and forecasts by developing a machine learning model of FIES-based food insecurity rates based on covariates that are available for territories where actual survey data are not available. This model is then employed to obtain forecasts to 2030 and assess likely future trends in food insecurity. We take advantage of individual-level characteristics collected in FAO's FIES surveys to push the analysis beyond the country level to a sub-national level. To our knowledge, this is the first effort to model FIES-based food insecurity at that scale.

@@

\section{Methods}
\subsection{Disaggregation}
The data on the FIES collected by Gallup on behalf of FAO also records many individual-level attributes from the respondents, including age, gender, wealth quintile, and whether they live in an urban or rural area.  Using national data on these variables, a standard weighting scheme is created for each individual, based on the ratio of population probabilities to sample probabilities of an individual with those characteristics being selected \citep{bethlehem2009applied}.  We use a similar methodology and calculate post-stratification weights at a subnational level.  We use year-specific subnational estimates of population shares by gender, age, urbanization, and, where available, wealth, and create a separate set of weights for each subnational area across all individuals.  For age and gender, we used gridded data from WorldPop \citep{Tatem2017}; for urbanization, we used spatially explicit estimates published by Jones and O'Neill \citep{Jones2016}; and for wealth we used data on household wealth quintiles from Demographic and Health Surveys \citep{dhsall}.  This methodology rests on the mild assumption that, within a given country, two individuals of the same age, gender, urbanization level and wealth quintile will have a similar food security status.

\subsection{Covariates}
We model subnational rates of moderate and severe food insecurity as a function of several covariates related to human development, nutrition, health, infrastructure, income levels and distribution, and the environment (See Table \ref{tab:covars}).  We employ covariates for which projections are available up to year 2030 and, in most cases, also have subnational spatial resolution.  For projections based on the Shared Socioeconomic Pathways (SSP) framework \citep{oneill2014new}, we use the middle-of-the-road pathway, SSP2, and for climatological variables, we use forecasts based on the Representative Concentration Pathway (RCP) 6.0 \citep{van2011representative}.

Many of the variables require the harmonization of subnational historical data with projections at the national level.  These include mean years of schooling, GDP per capita, as well as population. We use observed trajectories in the historical distribution of variables among subnational areas within a country to predict the future distribution of GDP, population, and schooling among subnational areas and disaggregate national-level future projections.

For health variables that have historically shown long-term trends and represent a rate of occurrence in a population, including stunting, wasting, and malaria mortality rate, we estimate the Annualized Rate of Change (AROC) for each subnational area to model these and to obtain predictions for variables up to 2030. We calculate the rate of change between each pair of years in the dataset, and then use the mean rate of change over the period for which data are available, giving greater weight to more recent years, and applying that mean rate of change to estimate future levels.  This method is used to obtain forecasts of the rates of stunting and wasting up to the year 2030.

@@

\subsection{Modeling}
We fit two models for all subnational areas globally from 2010 to 2030: one for the percent of the population over the threshold for moderate food insecurity, and one for the percent of the population over the threshold for severe food insecurity.  We use random forest regressions, which perform well in the presence of non-linearities and interaction effects among predictor variables \citep{hastie2009elements}.  A random forest regression involves generating a large number of decision trees, each using different sub-samples of the data, and then aggregating the predictions of the decision trees.  By taking an ensemble of decision tree models, random forests introduce more variance and balance out the bias that is common to methods based on single decision trees \citep{friedman2001elements}. Given the bounded nature of our variables of interest, the random forest regression are applied to the logistically transformed rates.  We create our models using the \texttt{randomForestSRC} package in R \citep{ishwaran2019randomforestsrc}.

@@

When performing cross validation, it is typical to select observations at random.  However, in our training data we have multiple highly similar observations from each country across years and subnational areas, and we extend our model predictions to countries that were not found at all in our training data.  Thus, we perform a type of leave-one-out cross validation, where we leave out all of the data from one country for each iteration.  Thus, the model error after cross validation gives us a sense of how the model is performing in countries where FIES data was not collected.  By this validation metric, we find that our models perform well, with an $R^2$ of 0.792 and mean average error (MAE) of 0.071 for moderate food insecurity and an $R^2$ of 0.746 and a MAE of 0.041 for severe food insecurity.  After fitting models on the entire dataset, we observe a within-sample $R^2$ of 0.950 and an MAE of 0.034 for moderate food insecurity and an $R^2$ of 0.945 and a MAE of 0.018 for severe food insecurity.

@@

\subsection{Food Insecurity Across Space}
We find substantial heterogeneity in the global distribution of severe food insecurity across countries and subnational regions (See Fig. \ref{fig:map}).  For the year 2020, mainland sub-Saharan Africa is the continent with the highest rates of severe food insecurity, with 20\% of people over the corresponding threshold in at least one subnational area in every country except Gabon, Djibouti and Equatorial Guinea.  Outside of Africa, serious pockets of severe food insecurity also occur in Venezuela, Syria, Papua New Guinea, Yemen, and Afghanistan.  In many large middle-income countries, severe food insecurity is also quite prevalent, with rates between 10\% and 15\% in northern Brazil, many central Asian and middle-eastern countries, as well as India and Indonesia.

@@

@@

\section{Discussion}
This paper presents the first global estimates of the experience of food insecurity at a subnational level, with forecasts until the year 2030.  We find that, while food insecurity has been increasing and has likely worsened in 2020 due to the economic downturn associated with the COVID-19 pandemic, our models predict that the percent of people who experience food insecurity globally will decrease throughout the 2020s, largely because of economic growth in developing countries.  Nevertheless, lower and lower-middle income countries are expected to see overall increases in the number of moderately food insecure people.

@@

\section{Additional Information}
All code necessary to reproduce these results is on Github at \href{https://github.com/mcooper/wdl-fies}{github.com/mcooper/wdl-fies}. We have no competing interests to disclose.

>>>>

%\maketitle
\begin{abstract}
This paper presents the first global picture of food security at an administrative area, subnational level based on the Food Insecurity Experience Scale, the indicator for the Sustainable Development Goal of ``Zero Hunger" that is most indicative of the individual's lived experience of food insecurity and hunger.  Using microdata from 75 countries and filling gaps using machine learning, we find significant heterogeneity in levels of food insecurity around the world.  Examining global temporal trends and accounting for the effects of the COVID-19 pandemic, we find that rates of severe food insecurity are declining, resulting in global decreases in the total number of severely food insecure people.  However, the total number of moderately food insecure people has been increasing and, after recovering from the shock of the COVID-19 pandemic, we predict it will continue to increase through the end of the 2020s.  Overall, we conclude that current trends in development and demographic change will still leave a large share of the world's population still experiencing hunger by 2030.  \end{abstract}

@@

This paper contributes to shifting the research frontier by filling these two gaps in terms of spatial coverage and forecasts by developing a machine learning model of FIES-based food insecurity rates based on covariates that are available for territories where actual survey data are not available. This model is then employed to obtain forecasts to 2030 and assess likely future trends in food insecurity. We take advantage of individual-level characteristics collected in FAO's FIES surveys to push the analysis beyond the country level to a sub-national level. To our knowledge, this is the first effort to model FIES-based food insecurity at that scale and make projections to 2030.

@@

\section{Methods}
\subsection{Disaggregation}
The data on the FIES collected by Gallup on behalf of FAO is collected annually within a country, and is therefore at the country-year level.  This data also records many individual-level attributes from the respondents, including age, gender, wealth quintile, and whether they live in an urban or rural area.  Using national data on these variables, a standard weighting scheme is created for each individual, based on the ratio of population probabilities to sample probabilities of an individual with those characteristics being selected \citep{bethlehem2009applied}.  In order to disaggregate the data to the administrative area level, we use a similar methodology and calculate post-stratification weights.  We use year-specific administrative area estimates of population shares by gender, age, urbanization, and, where available, wealth, and create a separate set of weights for each administrative area across all individuals.  For subnational distributions of age and gender, we rely on gridded population data.  While there are many different gridded population datasets, they often take different approaches to disaggregating population data, are available at different temporal resolutions, and have varying availability of demographic data \cite{Ley2019Sep, friedman2001elements, Tuholske2019Aug}.  We therefore chose to use the WorldPop dataset, which has demographic data on age and sex and is available annually \citep{Tatem2017}.  For rates of urbanization by administrative area, we used spatially explicit estimates published by Jones and O'Neill \citep{Jones2016}, and for wealth we used data on household wealth quintiles from Demographic and Health Surveys \citep{dhsall}.  This methodology rests on the mild assumption that, within a given country, two individuals of the same age, gender, urbanization level and wealth quintile will have a similar food security status.

\subsection{Covariates}
We model rates of moderate and severe food insecurity as a function of several covariates related to human development, nutrition, health, infrastructure, income levels and distribution, and the environment at the administrative area level (See Table \ref{tab:covars}).  We employ covariates for which projections are available up to year 2030 and, in most cases, also have administrative area spatial resolution.  For projections based on the Shared Socioeconomic Pathways (SSP) framework \citep{oneill2014new}, we use the middle-of-the-road pathway, SSP2, and for climatological variables, we use forecasts based on the Representative Concentration Pathway (RCP) 6.0 \citep{van2011representative}.

Many of the variables require the harmonization of subnational historical data with projections at the national level.  These include mean years of schooling, GDP per capita, as well as population. We use observed trajectories in the historical distribution of variables among administrative areas within a country to predict the future distribution of GDP, population, and schooling among these areas and disaggregate national-level future projections.

For health variables that have historically shown long-term trends and represent a rate of occurrence in a population, including stunting, wasting, and malaria mortality rate, we estimate the Annualized Rate of Change (AROC) for each administrative area to model these and to obtain predictions for variables up to 2030. We calculate the rate of change between each pair of years in the dataset, and then use the mean rate of change over the period for which data are available, giving greater weight to more recent years, and applying that mean rate of change to estimate future levels.  This method is used to obtain forecasts of the rates of stunting and wasting up to the year 2030.

@@

\subsection{Modeling}
We fit two models for all administrative areas globally from 2010 to 2030: one for the percent of the population over the threshold for moderate food insecurity, and one for the percent of the population over the threshold for severe food insecurity.  We use random forest regressions, which perform well in the presence of non-linearities and interaction effects among predictor variables \citep{hastie2009elements}.  A random forest regression involves generating a large number of decision trees, each using different sub-samples of the data, and then aggregating the predictions of the decision trees.  By taking an ensemble of decision tree models, random forests introduce more variance and balance out the bias that is common to methods based on single decision trees \citep{friedman2001elements}. Given the bounded nature of our variables of interest, the random forest regression are applied to the logistically transformed rates.  We create our models using the \texttt{randomForestSRC} package in R \citep{ishwaran2019randomforestsrc}.

@@

When performing cross validation, it is typical to select observations at random.  However, in our training data we have multiple highly similar observations from each country across years and administrative areas, and we extend our model predictions to countries that were not found at all in our training data.  Thus, we perform a type of leave-one-out cross validation, where we leave out all of the data from one country for each iteration.  Thus, the model error after cross validation gives us a sense of how the model is performing in countries where FIES data was not collected.  By this validation metric, we find that our models perform well, with an $R^2$ of 0.792 and mean average error (MAE) of 0.071 for moderate food insecurity and an $R^2$ of 0.746 and a MAE of 0.041 for severe food insecurity.  After fitting models on the entire dataset, we observe a within-sample $R^2$ of 0.950 and an MAE of 0.034 for moderate food insecurity and an $R^2$ of 0.945 and a MAE of 0.018 for severe food insecurity.

@@

\subsection{Food Insecurity Across Space}
We find substantial heterogeneity in the global distribution of severe food insecurity across countries and administrative areas (See Fig. \ref{fig:map}).  For the year 2020, mainland sub-Saharan Africa is the continent with the highest rates of severe food insecurity, with 20\% of people over the corresponding threshold in at least one administrative area in every country except Gabon, Djibouti and Equatorial Guinea.  Outside of Africa, serious pockets of severe food insecurity also occur in Venezuela, Syria, Papua New Guinea, Yemen, and Afghanistan.  In many large middle-income countries, severe food insecurity is also quite prevalent, with rates between 10\% and 15\% in northern Brazil, many central Asian and middle-eastern countries, as well as India and Indonesia.

@@

We also find substantial within-country variation in rates of moderate and severe food insecurity in many countries.  In Brazil, for example, northern regions are much more food insecure than southern regions, while in China, coastal regions and administrative areas that contain large cities such as Beijing and Shanghai are more food secure.  Finally, in many African countries, arid regions have higher rates of moderate and severe food insecurity, such as the Sahel in Mali and Niger, the Central and Western Zones of Tanzania, and the Somali Region of Ethiopia.

@@

\section{Discussion}
This paper presents the first global estimates of the experience of food insecurity at a administrative area level using the FIES, with forecasts until the year 2030.  We find that, while food insecurity has been increasing and has likely worsened in 2020 due to the economic downturn associated with the COVID-19 pandemic, our models predict that the percent of people who experience food insecurity globally will decrease throughout the 2020s, largely because of economic growth in developing countries.  Nevertheless, lower and lower-middle income countries are expected to see overall increases in the number of moderately food insecure people.

@@

\section{Competing Interests}
We have no competing interests to disclose.
